% arara: pdflatex: { synctex: yes }
% arara: makeindex: { style: ctuthesis }
% arara: bibtex

% The class takes all the key=value arguments that \ctusetup does,
% and a couple more: draft and oneside
\documentclass[twoside]{ctuthesis}

\ctusetup{
	preprint = \ctuverlog,
	mainlanguage = english,
	titlelanguage = english,
%	mainlanguage = czech,
	otherlanguages = {czech},
	title-czech = {Částečné učení s učitelem pro časově-prostorovou segmentaci satelitních snímků},
	title-english = {Semi-Supervised Learning for Spatio-Temporal Segmentation of Satellite Images},
	%subtitle-czech = {Cesta do tajů kdovíčeho},
	%subtitle-english = {Journey to the who-knows-what wondeland},
	doctype = M,
	faculty = F3,
 	department-czech = {Katedra kybernetika},
	department-english = {Department of Cybernetics},
	author = {Antonín Hruška},
	supervisor = {doc. Boris Flach, Dr. rer. nat. habil.},
	%supervisor-address = {ss},
	%supervisor-specialist = {John Doe},
	fieldofstudy-english = {Cybernetics and Robotics},
	%subfieldofstudy-english = {Mathematical Modelling},
	fieldofstudy-czech = {Kybernetika a robotika},
	%subfieldofstudy-czech = {Matematické modelování},
	keywords-czech = {Segmentace obrazu, SSL,HVAE,LVAE,symmetric learning, MixMatch,CityScape},
	keywords-english = {Image segmentation, SSL, VAE,HVAE, LVAE, symetric learning, MixMatch, CityScape},
	day = 26,
	month = 5,
	year = 2023,
	specification-file = {Thesis_Assignment_Antonin_Hruska_Semi-Supervised_Learning_for_Spatio-Temporal_Segmentation_of_Satellite_Images},
	front-specification = true,
	front-list-of-figures = true,
	front-list-of-tables = true,
    pkg-hyperref = true,
% Does not work	pkg-biblatex = true
%	monochrome = true,
%	layout-short = true,
}

\ctuprocess

\addto\ctucaptionsczech{%
	\def\supervisorname{Vedoucí}%
	\def\subfieldofstudyname{Studijní program}%
}

\ctutemplateset{maketitle twocolumn default}{
	\begin{twocolumnfrontmatterpage}
		\ctutemplate{twocolumn.thanks}
		\ctutemplate{twocolumn.declaration}
		\ctutemplate{twocolumn.abstract.in.titlelanguage}
		\ctutemplate{twocolumn.abstract.in.secondlanguage}
		\ctutemplate{twocolumn.tableofcontents}
		\ctutemplate{twocolumn.listoffigures}
	\end{twocolumnfrontmatterpage}
}

% Theorem declarations, this is the reasonable default, anybody can do what they wish.
% If you prefer theorems in italics rather than slanted, use \theoremstyle{plainit}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{note}
\newtheorem*{remark*}{Remark}
\newtheorem{remark}[theorem]{Remark}

\setlength{\parskip}{4ex plus 0.2ex minus 1ex}

% Abstract in Czech
\begin{abstract-czech}
Tato diplomová práce se zabývá využitím algoritmů částečného strojového učení s učitelem pro segmentaci
lesních ploch ze satelitních snímků. Místo přímého vyhodnocování na satelitních snímcích budeme používat 
veřejně dostupný dataset CityScape pro ověření výsledků, reprodukovatelnost a porovnání s jinými metodami. 

První část diplomové práce zahrnuje rešerši state-of-the-art algoritmů v oblasti částečného strojového učení s učitelem. Představujeme
dva vyhodnocované algoritmy: referenční MixMatch a nový algoritmus založený na symetrickém učení variačních autoenkoderů.
Poskytujeme také samostatný úvod do variačních autoenkoderů, zahrnující jejich teoretické základy a rozšíření, aby čtenář
porozuměl základním principům, na kterých nový algoritmus staví.

Hlavním cílem je adaptovat a posoudit schopnosti obou nového a referenčního algoritmu v oblasti segmentace. V obou případech
využíváme stejnou architekturu modelu založenou na U-net síti pro srovnatelnost a vyhodnocení.
\todo{Rewrite once results are obtained}
\end{abstract-czech}

% Abstract in English
\begin{abstract-english}
	This master's thesis explores the application of semi-supervised machine learning algorithms for segmenting 
	forest land cover in satellite images. Instead of directly evaluating these algorithms on satellite imagery, we employ
	the publicly available CityScape dataset to verify results, ensure reproducibility, and facilitate comparison with other 
	existing methods.

	The thesis begins with an in-depth review of the current state-of-the-art in semi-supervised machine learning algorithms. We 
	introduce two evaluated algorithms: MixMatch, which serves as a reference, and a novel algorithm based on symmetric learning of
	variational autoencoders. To provide readers with a comprehensive understanding, we also provide a self-contained introduction 
	to variational autoencoders, covering their theoretical foundations and extensions.
	
	The primary objective of this work is to adapt and evaluate the capabilities and segmentation potential of both the new algorithm
	and the reference algorithm. To ensure comparability, we employ a U-net network-based model architecture in both cases.
	\todo{Rewrite once results are obtained}
\end{abstract-english}

% Acknowledgements / Podekovani
\begin{thanks}
	I would like to express my deepest gratitude to Doc. Boris Flach, Dr. rer. nat. habil., without 
	whom this endeavor would not have been possible. I am immensely thankful to him for the professional 
	guidance he provided me with, as well as his kind words and moral support during the challenging times. 
	I would also like to extend my heartfelt thanks to all the teachers, lecturers, and individuals who generously
	 dedicated their time, effort, and passion to teach me new things and propel me forward. 

	 I am also grateful for the access to the computational infrastructure of the OP VVV funded project CZ.$02.1.01/0.0/0.0/16\_019/0000765$. 
	 The support provided by the "Research Center for Informatics" through this project is sincerely acknowledged and appreciated.

	 In addition, I would like to express my gratitude to my family, partner, and close friends who have been unwavering 
	 in their support throughout my studies and have played a significant role in helping me complete this thesis. Their
	encouragement and assistance have been invaluable to me, and I am truly grateful for their presence in my life.
\end{thanks}

% Declaration / Prohlaseni
\begin{declaration}
\label{sec:decleration}
I declare that the presented work was developed independently and that I have listed all sources of information
used within it in accordance with the methodical instructions for observing the ethical principles in the preparation 
of university theses.

In particular, I declare that I have not used natural language processing models to generate the ideas and themes used in the thesis. 
Nevertheless, I have used such tools (Grammarly, DeepL) to correct any grammatical or syntactical errors or to translate the original 
Czech version of the text into English and to improve the wording of such text. I have also used the Chat-GPT models for creating code
snippets in both Python and Latex, as well as having a second reader of the thesis. 

In Prague, 26th May, 2023
\end{declaration}

\usepackage{diagbox}
\usepackage{siunitx}
\sisetup{output-decimal-marker={.},exponent-product=\cdot}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{shapes, arrows.meta, positioning}
\usetikzlibrary{positioning,automata}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{makecell}
\usepackage{physics}
\usepackage{lipsum}
\usepackage{listings}
%\usepackage{julia-mono-listings}

\usepackage{algorithm}
\usepackage{algpseudocode}

% CUSTOM FONT SETTING
\usepackage{tgpagella}

\usepackage{xkeyval}	% Inline todonotes
\usepackage[textsize = footnotesize]{todonotes}
\presetkeys{todonotes}{inline}{}

\usepackage[style=numeric]{biblatex}
\addbibresource{Diploma Thesis.bib}
\DefineBibliographyStrings{english}{%
  bibliography = {References},
}

\usepackage{url}
\usepackage{tabularx}
\usepackage{bookmark}

%%% --- The following two lines are what needs to be added --- %%%
\setcounter{biburllcpenalty}{7000}
\setcounter{biburlucpenalty}{8000}

% Definition of new commands
\newcommand{\argmax}{\mathop{\rm argmax}}
\newcommand{\argmin}{\mathop{\rm argmin}}
\newcommand{\arctg}{\mathop{\rm arctg}}
\newcommand{\tg}{\mathop{\rm tg}}
\newcommand{\aff}{\mathop{\rm aff}}
\newcommand{\conv}{\mathop{\rm conv}}
%\newcommand{\rank}{\mathop{\rm rank}}
\newcommand{\diag}{\mathop{\rm diag}}
\newcommand{\sgn}{\mathop{\rm sgn}}
\newcommand{\Null}{\mathop{\rm null}}
\newcommand{\Rng}{\mathop{\rm rng}}
\newcommand{\dist}{\mathop{\rm dist}}
\renewcommand{\d}[1]{\mbox{\rm d}#1}

%\renewcommand{\phi}{\varphi}

% Change of symbols for footnotes
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\graphicspath{{./figs/}}
% Boldface vectors. Example: \_x
\def\_#1{\protect\mathchoice
    {\mbox{\boldmath $\displaystyle\bf#1$}}
    {\mbox{\boldmath $\textstyle\bf#1$}}
    {\mbox{\boldmath $\scriptstyle\bf#1$}}
    {\mbox{\boldmath $\scriptscriptstyle\bf#1$}}}

% Blackboard letters. Example: \bb R
\def\bb#1{\mathbb{#1}}

% Custom commands
\def\maxjerk{j_{\mathrm{M}}}

% Rozsireni tabulky kvuli nahnacani symbolu.
\renewcommand{\arraystretch}{1.25}

% Cpp, C#
\newcommand{\CC}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\tiny\bf +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\tiny\bf +}}
\newcommand{\CS}{C\nolinebreak\hspace{-.05em}\raisebox{.6ex}{\tiny\bf \#}}

% repeating footnote
\newcommand{\savefootnote}[2]{\footnote{\label{#1}#2}}
\newcommand{\repeatfootnote}[1]{\textsuperscript{\ref{#1}}}

% Colour definitions
\definecolor{myBlue}{HTML}{6666FF}
\definecolor{myGreen}{HTML}{66B366}
\definecolor{myRed}{HTML}{FF6666}

% Make sumint symbol


\DeclareRobustCommand{\bbone}{\text{\usefont{U}{bbold}{m}{n}1}}

\DeclareMathOperator{\EX}{\mathbb{E}}

\begin{document}

\maketitle


\chapter{Introduction}
\input{_introduction}

\chapter{State of the art of SSL}
\label{motivation-theory}
In this chapter, we will present a brief introduction to semi-supervised learning in section~\ref{sec:ssl-introduction}.
We will discuss the key concepts and principles underlying semi-supervised learning and its significance in machine learning. 
Furthermore, we will classify the algorithms used in semi-supervised learning into various groups based on the different ideas 
and paradigms they employ.

Next, we will delve deeper into two specific approaches: MixMatch and Variational Autoencoders (VAEs). Section~\ref{sec:mix-match} 
will provide an in-depth introduction to MixMatch, while section~\ref{sec:vae} will focus on the basics of Variational Autoencoders, 
introducing the topic, its achievements, and recognized shortcomings. Following that, in section~\ref{sec:hiearchical-vae}, we will
explore advanced variants of VAEs that partially overcome these shortcomings and represent the state-of-the-art in the field of VAE.

These sections serve as a foundation for comprehending symmetric equilibrium learning in VAEs, which will be introduced in section~\ref{sec:symmetric_learning}.
Additionally, section~\ref{sec:exp_family} provides an introduction to the family of exponential 
distributions, which are extensively used in VAE and its advanced variants.
\todo{Add reference to unpublished paper}

% SotA SSL
\input{_ssl}
% VAE, HVAE,LVAE
\input{_vae}
% Exp family 
\input{_exp_family}
% Symmetrical learning paper
\input{_sym_learning}

\chapter{Methods}
\label{methods}
% Metics, dataset, task etc.
\input{_problem_definition}
% Mixmatch adaptation
\input{_mixmatch_adaptation}
% HVAE design and choice 
\input{_sym_learning_adaptation}



\chapter{Experiments \& Results} 
\label{results}
\input{_experiments}

\todo{Presentation of the results of the experiments}
\todo{Comparison of the results of the ladder autovariational encoder with symmetrical learning and mixmatch models}
\todo{Discussion of the results and analysis of the findings}



\chapter{Conclusion}
\label{conclusions}
\todo{2-3 pages}	
\todo{Summary of the main findings and contributions}

\printbibliography

%\ctutemplate{specification.as.chapter}

\chapter*{List of used symbols, notations}
\input{_list_of_symbols}

\end{document}